Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-11 paper/disc20clustering.tex
Table width
Stretch
some changes to PPL model and k=2 comments

\begin{table*}[h]
	\centering
	\renewcommand{\arraystretch}{1.5}
	\begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
		\rowcolor{gray!50}
		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
		\rowcolor{gray!25}
		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
		\rowcolor{gray!25}
		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
	\end{tabular}
	\caption{Overview of known results and our contributions. The first three rows summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
	}
	\label{tab:overview}
	\vspace{-7mm}
\end{table*}

@@

We present an asymptotically optimal algorithm for the learning problem.
The algorithm immediately collocates every communicating pair and never separates them.
To help maintain the collocation of communicating pairs, we use the concept of components, introduced by Avin et al. \cite{repartition-disc}.

The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model, we choose the \maciek{component respecting?} partition closest to the initial partition.
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).

In our algorithms, a~\emph{component} is a subset of frequently communicating nodes.
We keep all nodes of each component collocated in the same cluster, similar to Avin et al.~\cite{repartition-disc}.
When we  move a~node, we move the whole component that contains it.
We maintain a balanced partition of our components,
a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.

We say that a component is \emph{trivial} if it contains exactly one node,
and the node in this component is \emph{isolated}.
We say that a~partition is \emph{component-respecting}
if the nodes belonging to the same component are always collocated.

@@

With each repartitioning,
\PPL moves to a \maciek{component-respecting} partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond the distance $\Delta^*$.

@@

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final partition $P_F$
	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all \maciek{component-respecting} $\ell$-way partitions of components
	that are in the same (closest) distance to $P_I$.
	That is, 
	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
	it does not move to a partition
	$P', \Delta(P') > \Delta^*$,
	before it enumerates all partitions at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
	With each distinct request, 
	the size of some component increases by one.
	For any cluster $F_i \in P_F$,
	we have $\sum_{C \in F_i} |C| = k$.
	A component $C \in F_i$,
	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
	Hence, the total number of times a component in $F_i$ grows is 
	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
	for which \PPL performs a repartitioning and
	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
	By Lemma \ref{lemma:rebalancecost},
	each repartitioning costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
\end{proof}

@@

We charge both \OPT and \RM whenever \RM collocates a pair.%(i.e., on a collocation event).
\RM collocates a pair always with a swap,%(i.e., two simultaneous migrations),
which  costs $2\alpha$,
while OPT may save some costs by collocating multiple pairs at once.
% before they inflict too much communication cost.
 Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
Hence,
whenever \OPT collocates a pair,
we charge it only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).

@@

\end{proof}
\maciek{Fonts of OPT and \OPT differ. The latter is better.}

\maciek{Excessive comma before ", is $\lambda +2\alpha$?}

@@

\maciek{"then \OPT pays $\lambda$ for serving" comma before then?}

\maciek{", or $\alpha$ for collocating \textbf{and serving some requests remotely}" remove what's bold? The migration is sufficient}

\maciek{"it can have one edge already collocated" at most one edge?}

\maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call. Maybe use the same formulation? I feel like one sentence can be enough to explain this}

\maciek{Suggestion: maybe call "its intervals" "\textbf{the saturating intervals}"?}

\maciek{OPT's cost formula: make it one line by removing the expression with minimums. Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula)}

\maciek{"	I.e., $ C^t_{\set{u,v}} = \lambda$." is too short for a sentence. Merge with the previous sentence?}

\maciek{"and for final costs at t=m" for final requests}

\maciek{"then the total cost to OPT" cost for OPT}

\maciek{(strong) Suggestion: instead of "this means", "meaning", I saw exclusively the use of "therefore", "hence" and "thus", and maybe "this implies" in papers. Also in PPl}

@@

\mahmoud{mention the offline approx.}

>>>>

\begin{table*}[h]
	\centering
	\renewcommand{\arraystretch}{1.5}
	\begin{tabular}{>{\centering\arraybackslash}p{4.25cm}|>{\centering\arraybackslash}p{4.25cm}>{\centering\arraybackslash}p{4.25cm}}
		\rowcolor{gray!50}
		\textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
		\textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
		\rowcolor{gray!25}
		\textbf{$k=3$}&  $\Omega(\ell)$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $O(\ell) $\hspace{0.3cm}(\S \ref{sec:k3})\\
		$k > 3$ & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k^2 \cdot \ell^2)$\hspace{0.1cm} \cite{repartition-disc} \\
		\rowcolor{gray!25}
		Learning model & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell)$\hspace{0.1cm} (\S \ref{sec:ppl}) \\
	\end{tabular}
	\caption{Overview of known results and our contributions. The first three rows summarizes the results for the general partitioning model, except for the last row that summarizes the results for the learning model for arbitrary $k$ and $\ell$.
	}
	\label{tab:overview}
	\vspace{-7mm}
\end{table*}

@@

We present an asymptotically optimal algorithm for the learning problem.
The algorithm  collocates  a pair as soon as they communicate and it never separates them.
In order to preserve collocated pairs,
we employ the concept of components,
introduced by Avin et al. \cite{repartition-disc}.

%\medskip
%In our algorithms, 
A~\emph{component} is a subset of frequently communicating nodes.
We keep all nodes of a component always collocated in the same cluster,
i.e., when we  move a~node,
we move the whole component that contains it.
A~partition that maintains this invariant is a \emph{component-respecting} partition.
%We say that a~partition is \emph{component-respecting}
%if the nodes belonging to the same component are always collocated.
We say that a component is \emph{trivial} if it contains exactly one node,
and the node in this component is \emph{isolated}.

In addition,
we maintain a balanced partition of our components as long as such partition exists,
a reminiscent of partitioning given integers into sets of equal sum \cite{integer-partitions-book}.
In contrast, our partition is time-varying:  two components are merged into one component once they communicate, and we adjust the partition accordingly.

The algorithm in this section and the algorithm for $k=3$ (cf. Section~\ref{sec:k3}) are modified versions of the algorithm DET from \cite{repartition-disc}.
The difference is in the choice of partition after a component merge.
In DET, the partition was arbitrary.
In the algorithm for the learning model,
we choose a component-respecting partition closest to the initial partition.
In the algorithm from Section~\ref{sec:k3}, we choose the partition closest to the current partition (a repartition of minimum cost).
A generic sketch of these two algorithms can be found in the appendix (Algorithm \ref{alg:ppl}).

@@

With each repartitioning,
\PPL moves to a component-respecting partition that minimizes the distance to the initial partition $P_I$.
As a result,
\PPL never moves to a~partition that is more than $\Delta^*$  node migrations away from $P_I$.
This invariant latter ensures us that \PPL does not pay too much while recovering $P_F$.
We emphasize that a repartitioning by \PPL replaces the current partition $P$ with a~perfect partition closest to $P_I$.
Meaning, it never moves to a partition beyond the distance $\Delta^*$.

@@

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final partition $P_F$
	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all component-respecting $\ell$-way partitions of components
	that are in the same (closest) distance to $P_I$.
	That is, 
	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
	it does not move to a partition
	$P', \Delta(P') > \Delta^*$,
	before it enumerates all partitions at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
	With each distinct request, 
	the size of some component increases by one.
	For any cluster $F_i \in P_F$,
	we have $\sum_{C \in F_i} |C| = k$.
	A component $C \in F_i$,
	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
	Hence, the total number of times a component in $F_i$ grows is 
	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
	for which \PPL performs a repartitioning and
	 \PPL performs at most $(k-1)\cdot\ell $ repartitions.
	By Lemma \ref{lemma:rebalancecost},
	each repartitioning costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
\end{proof}

@@

We charge both \OPT and \RM whenever \RM collocates a pair.%(i.e., on a collocation event).
\RM collocates a pair always with a swap,%(i.e., two simultaneous migrations),
which  costs $2\alpha$,
while \OPT may save some costs by collocating multiple pairs at once.
% before they inflict too much communication cost.
 Therefore it pays the price of only one migration per pair  (\mahmoud{see Figure \ref{fig:TBD}}).
Hence,
whenever \OPT collocates a pair,
we charge it only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ on \RM).

@@

\end{proof}

@@

\maciek{Suggestion: "Note that \OPT cannot have both..." (till the end of paragraph) You told me this in one sentence during the call. Maybe use the same formulation? I feel like one sentence can be enough to explain this \mahmoud{I removed one redundant sentence. Since this is the core of the proof let's keep it wordy.}}

\maciek{OPT's cost formula: make it one line by removing the expression with minimums. Move minimums to the paragraph text, by immediately expanding minimums (two steps in paragraph, instead of in the formula) \mahmoud{it is useful for reader's intuition to have the min till the end. So that they can remember why choice of $\lambda$ is not trivial.}}

@@
